Distributed Topology Discovery in Self-Assembled Nano Network-On-Chip

### 2014-07-08 trunk/PAPER/disr.tex
fixes for review

\label{sec:disr_concepts}
The main concept behind the kind of turn prohibition we want to
achieve is the \emph{segment}. A segment $S_y$ is basically a path of consecutive
nodes and links, starting with a link attached to a node belonging to
a different segment $S_x$ and ending with a link attached to a
node of another segment $S_z$. In other words, a segment is
a path connecting two other different segments. The example
in Figure~\ref{fig:segments} shows the segment with id $1.4$, starting
with a link connected to the node $9$, traversing nodes
$4$, $3$, $2$ and ending with the link connected to the node $1$. Other
segments depicted are, for example, $9.2$ (link from $9$, node $10$, link to node $3$) and $9.3$
(link from $9$, node $8$, link, node $7$, link, node $6$, link, node $5$, link to node $1$). Note
also that start/ending links can coincide and a segment can also be
formed by one single link (e.g. the case of $7.2$ and $5.4$).
An exception to the rule that a segment must connect two other
segments is the first segment established in the network, called
\emph{starting segment} which is a loop beginning/ending on the same
node.

\begin{figure}
\centering
    \includegraphics[width=0.25\textwidth]{pictures/network_slice.eps}
  \caption{Example of Segments}
  \label{fig:segments}
\end{figure}

A detailed description of the \disr{} execution model as depicted in
Figure~\ref{fig:disr_events} will be discussed in
Subsection~\ref{sub:phases}. Roughly speaking, the execution phases consist of:
\begin{enumerate}
\item Injection of the \disr{} process from an upper layer to set a \emph{bootstrap
node}.
\item Broadcasting and confirmation to create the first segment of the subnet.
\item Parallel requests starting from assigned node to discover other
segments. 
\item Turn prohibition in each confirmed segment.
\item Draining of remaining packet due to timeouts.
\end{enumerate}

@@

\begin{figure}
  \centering
    \includegraphics[width=0.25\textwidth]{pictures/dbs_updated.eps}
  \caption{\disr{} node execution model}
  \label{fig:state_machine}
\end{figure}

\begin{figure*}
\centering
\begin{tabular}{cccc}
\includegraphics[width=0.23\textwidth]{pictures/seq01.eps} & 
\includegraphics[width=0.23\textwidth]{pictures/seq02.eps} &
\includegraphics[width=0.23\textwidth]{pictures/seq03.eps} & 
\includegraphics[width=0.23\textwidth]{pictures/seq04.eps} \\
(a) & (b) & (c) & (d) \\
\includegraphics[width=0.23\textwidth]{pictures/seq05.eps} &
\includegraphics[width=0.23\textwidth]{pictures/seq06.eps} &
\includegraphics[width=0.23\textwidth]{pictures/seq07.eps} & 
\includegraphics[width=0.23\textwidth]{pictures/seq08.eps} \\
(e) & (f) & (g) & (h) \\
\end{tabular}
\caption{Example of \disr{} segment setup: Nodes' status is labeled as \emph{CandidateStarting} (\emph{Cs}), \emph{Candidate} (\emph{C}), \emph{Free} (\emph{F}), \emph{Assigned} (\emph{A}) or \emph{Bootstrap} (\emph{B})}
\label{fig:disr_events}
\end{figure*}

@@

\item{\textbf{Injecting Segment Requests}}: Each node in the
\emph{Assigned} status can potentially initiate a new segment request.
So it check its links and if a free link is found it changes its
status to \emph{ActiveSearching} \emph{(c)} and then sends
a \texttt{SEGMENT\_REQUEST} across the free link \emph{(c)} (see
Figure~\ref{fig:disr_events}-e). 
Note that each \emph{Assigned} node can inject the request just after
becoming \emph{Assigned}, i.e. it is not needed that the whole confirmation
process in completed (Figure~\ref{fig:disr_events}-f).
Note also that nodes previously set as \emph{CandidateStarting}, when receive
a \texttt{SEGMENT\_REQUEST} can simply cancel their
status and set themselves as \emph{Candidate} for that request, since this new
request means that the first starting segment has already been found \emph{(h)}. 
\item{\textbf{Segment Confirmation}}: The segment searching process is successful
when an already \emph{Assigned} node receives the
\texttt{SEGMENT\_REQUEST} packet. Then, a \texttt{SEGMENT\_CONFIRM} is
sent back along the path that originated the request
(Figure~\ref{fig:disr_events}-g), while the
node remains \emph{Assigned} (since it could confirm other segments). Each node
previously set as \emph{Candidate} for that segment id, when receives a confirmation packet changes its status to \emph{Assigned} \emph{(m)} and send back
the same confirmation until it reaches the initiator of the request,
which then changes from \emph{ActiveSearching} to \emph{Assigned} \emph{(d)}.

@@

\subsection{Intra-node vs Inter-node parallelism}
In the processes described above, we assumed that each node in the
\emph{Assigned} status can start a segment searching process by injecting a
\texttt{SEGMENT\_REQUEST}. However, a more accurate decision when
defining \disr{} should be made in order to decide whether or not the
node must actually perform this action. The point above is
strictly related to a general question: which level of
parallelism should we allow in \disr{}? The adopted approach, although
the nature of \disr{} itself is intrinsically parallel, is that the use of
parallelism makes things working in a more complex way. In other words,
\disr{} is parallel when needed, but it does not exploit parallelism
as an ``improving feature''. When not needed, things should be
serialized. For example: when a node is ``ready to search'' it could start
several segment searching processes associated to the same segment, one for
each free link. But serializing this operation, by investigating the free links
in order, could be a simpler solution. We refer to this saying that we
avoid intra-node parallelism.  Note that while intra-node parallelism
can be avoided, trying to avoid inter-node parallelism could be more
complex than allowing the parallelism itself. Imagine for example the
effort generated trying to coordinate nodes so that an unique segment searching
process is actually running in the whole subnet. Thus, in contrast to
the intra-node parallelism, the inter-node parallelism is a structural
property of the \disr{} algorithm and should not be avoided.
%------------------------------------------------------------------------------

>>>>

\label{sec:disr_concepts}
The main concept behind the kind of turn prohibition we want to
achieve is the \emph{segment}. A segment $S_b$ is basically a path of consecutive
nodes and links, starting with a link attached to a node belonging to
a different segment $S_a$ and ending with a link attached to a
node of another segment $S_c$. In other words, a segment is
a path connecting two other different segments. The example
in Figure~\ref{fig:segments} shows the segment with id $1.4$, starting
with a link connected to the node $9$, traversing nodes
$4$, $3$, $2$ and ending with the link connected to the node $1$. Other
segments depicted are, for example, $9.2$ (link from $9$, node $10$, link to node $3$) and $9.3$
(link from $9$, node $8$, link, node $7$, link, node $6$, link, node $5$, link to node $1$). Note
also that start/ending links can coincide and a segment can also be
formed by one single link (e.g. the case of $7.2$ and $5.4$).
An exception to the rule that a segment must connect two other
segments is the first segment established in the network, called
\emph{starting segment} which is a loop beginning/ending on the same
node.

\begin{figure}
\centering
    \includegraphics[width=0.45\textwidth]{pictures/network_slice.eps}
  \caption{Example of Segments}
  \label{fig:segments}
\end{figure}

A detailed description of the \disr{} execution model as depicted in
Figure~\ref{fig:disr_events} will be discussed in
Subsection~\ref{sub:phases}. Roughly speaking, the execution phases consist of:
\begin{enumerate}
\item Injection of the \disr{} process from an upper layer to set a \emph{bootstrap
node}.
\item Broadcasting and confirmation to create the first segment of the subnet.
\item Parallel segment requests starting from assigned nodes to discover other
segments. 
\item Resulting segment confirmation/cancel packets are processed.
\item Turn prohibition in each confirmed segment.
\end{enumerate}

This first work is explicitly focused on demostrating
the feasibility and scalability of \disr{} in the particular scenario
assumed; for this reason we will not investigate the further degree of
freedom that the last of the above steps introduces, i.e. choosing
\emph{where} actually placing the restriction in the context of the
segment. This choice would striclty depend on the particular routing
strategy that will be used and could result in some interesting
improvements. However, the scope of this work is understading how \disr{} can
implement a distributed topology discovery, mapping defects and
creating a segment partitioning that will be exploited in a second
design phase which includes choosing and implementing a routing strategy. For this
reason, the last two phases will be addressed as one, since we can
safely assume that a random turn (e.g. the first) will be phohibited
once each segment has been confirmed.

@@

\begin{figure}
  \centering
    \includegraphics[width=0.45\textwidth]{pictures/dbs_updated.eps}
  \caption{\disr{} node execution model}
  \label{fig:state_machine}
\end{figure}

\begin{figure*}
\centering
\begin{tabular}{cccc}
\includegraphics[width=0.23\textwidth]{pictures/seq01.eps} & 
\includegraphics[width=0.23\textwidth]{pictures/seq02.eps} &
\includegraphics[width=0.23\textwidth]{pictures/seq03.eps} & 
\includegraphics[width=0.23\textwidth]{pictures/seq04.eps} \\
(a) & (b) & (c) & (d) \\
\includegraphics[width=0.23\textwidth]{pictures/seq05.eps} &
\includegraphics[width=0.23\textwidth]{pictures/seq06.eps} &
\includegraphics[width=0.23\textwidth]{pictures/seq07.eps} & 
\includegraphics[width=0.23\textwidth]{pictures/seq08.eps} \\
(e) & (f) & (g) & (h) \\
\end{tabular}
\caption{Example of \disr{} segment setup: Nodes status is labeled as \emph{CandidateStarting} (\emph{Cs}), \emph{Candidate} (\emph{C}), \emph{Free} (\emph{F}), \emph{Assigned} (\emph{A}) or \emph{Bootstrap} (\emph{B})}
\label{fig:disr_events}
\end{figure*}

@@

\item{\textbf{Injecting Segment Requests}}: Each node in the
\emph{Assigned} status can potentially initiate a new segment request.
So it check its links and if a free link is found it changes its
status to \emph{ActiveSearching} \emph{(c)} and then sends
a \texttt{SEGMENT\_REQUEST} across the free link \emph{(c)} (see
Figure~\ref{fig:disr_events}-e). 
Note that each \emph{Assigned} node can inject the request just after
becoming \emph{Assigned}, i.e. it is not needed that the whole confirmation
process in completed (Figure~\ref{fig:disr_events}-f).
Note also that nodes previously set as \emph{CandidateStarting}, when receive
a \texttt{SEGMENT\_REQUEST} can simply cancel their
status and set themselves as \emph{Candidate} for that request, since this new
request means that the first starting segment has already been found \emph{(h)}.

\item{\textbf{Segment Confirmation}}: The segment searching process is
successful when an already \emph{Assigned} node receives the
\texttt{SEGMENT\_REQUEST} packet. Then, a \texttt{SEGMENT\_CONFIRM} is
sent back along the path that originated the request
(Figure~\ref{fig:disr_events}-g), while the node remains
\emph{Assigned} (since it could confirm other segments). Each node
previously set as \emph{Candidate} for that segment id, when receives
a confirmation packet changes its status to \emph{Assigned} \emph{(m)}
and send back the same confirmation until it reaches the initiator of
the request, which then changes from \emph{ActiveSearching} to
\emph{Assigned} \emph{(d)}.

@@

\subsection{Intra-node vs Inter-node parallelism}
In the processes described above, we assumed that each node in the
\emph{Assigned} status can start a segment searching process by
injecting a \texttt{SEGMENT\_REQUEST}. However, different choices
could have been made when defining \disr{}, strictly related to which
kind and level of parallelism should be supported. The approach adopted
currently is that, although the nature of \disr{} is intrinsically
parallel, the use of parallelism should not make things work in a too
complex/uncontrollable way. In other words, \disr{} is parallel when
needed, but it does not exploit parallelism as an ``improving
feature''. Thus, when not needed, things should be serialized. For
example: a node with free links could start several segment requests
associated to the same \emph{segID}, one for each free link. But
serializing this operation, by investigating the free links in order,
could be a simpler solution. We refer to this \disr{} design choice
saying that we avoid \emph{intra-node parallelism}.  Note that,
conversely, avoiding \emph{inter-node parallelism} could be more
complex than allowing it: for example, we can imagine the effort
generated when trying to coordinate nodes so that a unique segment
searching process is actually running in the whole subnet. Thus, in
contrast to the \emph{intra-node parallelism}, the \emph{inter-node
parallelism} is a structural property of the \disr{} algorithm and
should not be avoided.
%------------------------------------------------------------------------------

### 2014-07-08 trunk/PAPER/implementation.tex
fixes for review

A possible hardware implementation of \disr{} algorithm will be
described in this section. Figure~\ref{fig:node_structure} depicts the general
structure of a node in the particular scenario of DNA nano Network-On-Chip. In particular such a node
is composed by the following fundamental elements:

@@

To give a basic estimate of the overhead needed we focus on \disr{}-
specific components namely the control logic and configuration
registers needed to implement \disr{}. Figure~\ref{fig:implementation}
shows an architecture sketch, which mainly consist in the
following building elements:

@@

	\item \textbf{Packager:}
	Essentially it is a set of registers and multiplexers. Starting from
	an incoming packet stored in the input buffer, it updates packet's
	content before retransmission to a destination
	node.  In some cases, this circuitry builds the packet from
	scratch e.g. when a node is bootstrap and for
	the first time it should inject a \emph{starting\_segment\_request}. 
	Figure~\ref{fig:packager} depicts the architecture of this
	block. When a packet is created for the first time, the multiplexers 
	driven by the controller, will select information incoming from an internal 
    node such as the \emph{source\_ID}. In this situation the 
	TTL field should be reset to its initial value and the \emph{segment\_ID}
	is composed by the node ID and by the output port identification 
	number.

@@

All the blocks described above, reacts with the rising edge of
the \emph{clock signal}. An asynchronous system \emph{reset} signal is
also present to restore all registers to their default values. When
this signal is set, DBS state machines for each network
node are set as \emph{Free}.  
Figure~\ref{fig:pipeline} is shows the pipeline structure of the
\disr{} architecture. Packets incoming from neighbor nodes (or
generated locally) are sent to the DBS state machine in order to
ensure the status commutation. While DBS updates its state, in the
next clock cycle the control circuitry will update LED registers status and
will drive the \emph{Switch Matrix} in order to send eventually the packets to
the right output buffer.
%--------------------------------------------------------------------
\subsection{Synthesis results: Timing, Area and Power consumption considerations}

>>>>

A possible hardware implementation of \disr{} algorithm will be
described in this section. Figure~\ref{fig:node_structure} depicts the general
structure of a node in the particular scenario of DNA nano Network-On-Chip. In particular such a node
is composed of the following fundamental elements:

@@

To give a basic estimate of the overhead needed we focus on \disr{}-
specific components namely the control logic and configuration
registers needed to implement \disr{}. Figure~\ref{fig:implementation}
shows an architecture sketch, which mainly consists in the
following building elements:

@@

	\item \textbf{Packager:}
	Essentially it is a set of registers and multiplexers. Starting from
	an incoming packet stored in the input buffer, it updates packet's
	content before retransmission to a destination
	node.  In some cases, this circuitry builds the packet from
	scratch e.g. when a node is bootstrap and for
	the first time it should inject a \emph{starting\_segment\_request}. 
	Figure~\ref{fig:packager} depicts the architecture of this
	block. When a packet is created for the first time, the multiplexers 
	driven by the controller, will select information incoming from an internal 
    node such as the \emph{source\_ID}. In this situation the 
	TTL field should be reset to its initial value and the \emph{segment\_ID}
	is composed of the node ID and by the output port identification 
	number.

@@

All the blocks described above, reacts with the rising edge of
the \emph{clock signal}. An asynchronous system \emph{reset} signal is
also present to restore all registers to their default values. When
this signal is set, DBS state machines for each network
node are set as \emph{Free}.  
Figure~\ref{fig:pipeline} shows the pipeline structure of the
\disr{} architecture. Packets incoming from neighbor nodes (or
generated locally) are sent to the DBS state machine in order to
ensure the status commutation. While DBS updates its state, in the
next clock cycle the control circuitry will update LED registers status and
will drive the \emph{Switch Matrix} in order to send eventually the packets to
the right output buffer.
%--------------------------------------------------------------------
\subsection{Synthesis results: Timing, Area and Power consumption considerations}

### 2014-07-08 trunk/PAPER/introduction.tex
fixes for review

\section{Introduction and Motivation}

Exploring long term alternatives to the CMOS technology is gaining
more and more revelance as the scaling trend of such devices keeps
introducing new challenges: power density, defect tolerance, testing
costs and wire delays are only a few of the many critical aspects
involved~\cite{itrs13}. 
While software parallelism and multicore
approaches~\cite{horowitz2004} and \cite{powell2009} are partially mitigating the impact of such
constraints on performances, it is likely that the growing computing demand will
eventually need even more radical architectural modifications and new
paradigms in order to address the Computer Design challenges of the next
decades.

In the last years, Self-assembled nanoscale architectures~\cite{yan2003}
emerged as a promising technology due their tera/peta scale of
integration, defect tolerance and huge potential computing
capabilities. These technologies are certainly still at
their early stage of development, however different laboratory demos and
prof-of-concepts architectures have been presented~\cite{patwardhan2004, patwardhan2006_1, pistol2009}.
%, patwardhan2006
The main idea behind this approach is to exploit the physical regularity and
stability of DNA structures in order to create a scaffold onto which
nano-devices (e.g. nanowires and CNFETs~\cite{bachtold2001, tans1998, cui2001}) can be
attached. This can be achieved by designing appropriate complementary DNA tags for
each terminal to be placed, so that a nano device will be attached
only where its own DNA tag matches a complementary tag on the DNA grid
scaffold (see Figure~\ref{fig:nana}-a).
A detailed description of the chemical properties involved is far
beyond the scope of this paper (see also~\cite{braun1998, seeman1999}), so we will focus on
the three main challenges that this new fabrication process introduces
in Computer Design: \emph{(i) limited node complexity}, \emph{(ii) large scale
randomness} and \emph{(iii) high defect rates}.

The \emph{limited node complexity} aspect is directly
related to the use of complementary DNA tags in order to place circuit
components. In a traditional CMOS process the complexity is introduced
using a photolithography mask, so that larger (more complex) circuits
only require larger masks; conversely, self-assembly achieves
complexity by increasing the number of unique DNA tags, since more
different tags means having more control on component placement. Ideally, by specifying a single
and unique tag for each nano transistor terminal, we could exactly
choose where each component would be placed. But the number of DNA
symbols forming the DNA sequence is limited (sequence of 4 nucleobases
G,A,T,C) and so creating many different tags (of a given lenght) would
mean make them more similar to each other, increasing the probability
of incorrect/partial matching. To avoid this problem we should limit
the number of unique tags, thus limiting the complexity at each node.
In particular, a budget of about $10,000$ CNFETs per node has been
estimated in~\cite{liu_jetcs}.

\emph{Large scale randomness} is the other fundamental condition of
self-assembled technology: DNA tags allow a control placement inside
node grid, but there is no control of where each of these  grids will
be placed on the whole network. As a consequence, other typical
properties of regular networks cannot be guaranteed, e.g. being
connected to a fixed number of neighbors, having a determined
orientation and so on.

Finally, \emph{defect rates}: while they are very hardly tolerated in mask
based top down design , the same nature of a bottom up self
assembled process cannot assume such a deterministic device placement
process, thus defect tolerance is something like a design requirement
more than an exception to be avoided.

\begin{figure*}
\centering
\begin{tabular}{cc}
    \includegraphics[width=0.40\textwidth]{pictures/dna2b.eps} &
    \includegraphics[width=0.40\textwidth]{pictures/dna1_complex2.eps} \\
 (a) & (b)
 \end{tabular}
  \caption{(a) DNA tag matching and (b) resulting trregualr topology}
  \label{fig:nana}
\end{figure*}
These aspects of a DNA-self assembly process lead to some important
implications to be addressed when approaching to the design of
a computer architecture: computational model should be based on a
distributed network of small processing and storage nodes, randomly
placed and interconnected (see Figure~\ref{fig:nana}-b).
Proof-of-concept of such architectures can be found for
example in~\cite{patwardhan2006_1}, where instruction and data
operands are move around the network in order to be processed.
Indepentedly from the particular policy chosen for routing packets,
such networks must be able to guarantee deadlock freedom without any
prior assumption on topology.

In this work we introduce \disr{}, a Distributed Segment-based approach
to deadlock freedom in large scale DNA self assembled networks. Our
contribution aims to achieve the classical properties of a
segment-based approach~\cite{mejia_ipdps06} without requiring any
topology graph, external defect map or centralized algorithm
execution.  So the \disr{} approach is not intended to discover the
``optimal'' segment choice (ideally reachable with the knowledge of
the topology graph) but just to demonstrate a concrete model that can
fit into such complex, irregular and large sized networks.
It is important to
underline that, altougth we can assume an architecture similar to the
one depicted in Figure~\ref{fig:nana} as background, the topology discovery
problem faced in this work is more general and not strictly dependent on any of the
underlying computing architecture.

The paper is organized as follows: in Section~\ref{sec:related_works}
we summarize the main approaches to topology agnostic routing. Next in
Section~\ref{sec:disr_concepts} are described the main elements and
the execution model of the proposed approach.  In Section~\ref{sec:simulation}
simulations are carried out with the opensource tool Nanoxim to
demonstrate how \disr{} preserves some of the main segment-based
properties and how compares to the tree-based approaches. Finally a
draft of a architectural implementation is shown in
Section~\ref{sec:implementation} to evaluate the feasibility of the
approach in terms of node complexity.

\section{Background and Contribution}
\label{sec:related_works}
Several worksaddressed the problem of topology agnostic deadlock
freedom introducing virtual channels~\cite{sancho2002, skeie2002,
skeie2004, koibuchi2003}, but while they show interesting
performances, the low resources deadlock makes them not suitable for
the specific DNA self assembled networks we are assuming as
background.
Other topology independent algorithms~\cite{schroeder1991,
koibuchi2001, cherkasova1996} do not require any virtual channels but
exploit the concept of ``turn prohibition''.
In particular, authors of~\cite{Patwardhan05evaluatingthe} exploits the creation of a spanning tree of the
topology, placing bidirectional restrictions by avoiding a packet
to traverse the same link in both up and down directions.
The hierarchical nature of this approach can lead to uneven traffic
distribution, with many packets traversing upper links (near to the
root), but this is quite acceptable in classical wide area networks
topologies with a limited number of nodes. Other approaches such as
FX~\cite{sancho2000} mitigate this
issue, but the set of turn restrictions is still prefixed,
strictly depending on the particular tree root selected. 
Other solutions try to approach the issue of irregular
topologies by limiting the number~\cite{duato1997, gomez2004, koibuchi2008} or the
location~\cite{zhang2008, flich2008, liu2011} of missing links. This
restriction is clearly unacceptable given the hight-defect rates of
DNA self-assembled networks scenario. For the same reason, we also
avoid considering solutions based on hardware-redundancy to
dynamically recover defects as in~\cite{constantinides2006, kohler2010,  park2006, ebrahimi2013}.

In~\cite{mejia_ipdps06} authors present an approach the solves these
limitations by allowing turn restrictions to be placed locally,
independently from other restrictions. The whole network is
partitioned into segments, and a bidirectional turn
restriction can be freely chosen within a segment in order to guarantee
deadlock freedom while preserving network connectivity. This \emph{locality
independence} property, avoiding the requirement of any particular tree/root
choice, would make it the best choice for the given scenario;
however, it still requires the knowledge of the
whole network graph in order to find the segments.

The \disr{} approach presented in this work is a first attempt of
achieving the same goals of a segment based turn prohibition  without
requiring any topology graph or centralized execution. In particular
\disr{} implements segment discovery and turn prohibition using a
completely different mechanism, based on a distributed exchange of
small setup packets requiring on average $15\%$ of the available
resource budget. The following main features distinguish \disr{}
from classic approaches to segment partitioning: 
\begin{itemize}
\item No centralized entity is globally responsible and/or aware of
what is going on, i.e. the status of the \disr{} execution is
collectively distributed among the nodes
\item  No defect map and/or topology graph is used as input,
thus, the topology has to be discovered \emph{while} segments are
created
\item At the end of the execution, no segment list is created or
stored anywhere: each node is only aware of belonging to some segment,
ignoring the presence of other nodes in the same segment and even the
presence of other segments in the network
\end{itemize}

>>>>

\section{Introduction and Motivation}
\label{sec:introduction}

Exploring long term alternatives to the CMOS technology is gaining
more and more relevance as the scaling trend of such devices keeps
introducing new challenges: power density, defect tolerance, testing
costs and wire delays are only a few of the many critical aspects
involved~\cite{itrs13}. 
While software parallelism and multicore
approaches~\cite{horowitz2004,powell2009} are partially mitigating the impact of such
constraints on performances, it is likely that a growing computing demand will
eventually need even more radical architectural modifications and 
paradigms to address the Computer Design challenges of the next
decades.

In the last years, self-assembled nanoscale
architectures~\cite{yan2003} emerged as a promising technology due
their tera/peta scale of integration, defect tolerance and huge
potential computing capabilities. These technologies are certainly
still at their early stage of development, however different
laboratory demos and proof-of-concepts have been
presented~\cite{patwardhan2004, patwardhan2006, patwardhan2006_1,
pistol2009}. For a complete survey on self-assembled architectures and
how they compare against classical many-core systems in terms of
 technologies, fault tolerance, performance and software tools see
 also~\cite{Future}.
%, patwardhan2006
The main idea behind this approach is exploiting the physical
regularity and stability of DNA structures in order to create a
scaffold onto which nano-devices (e.g. nanowires and
CNFETs~\cite{bachtold2001, tans1998, cui2001}) can be attached. This
can be achieved by designing appropriate complementary DNA tags for
each terminal to be placed, so that a nano device will be attached
only where its own DNA tag matches a complementary tag on the DNA grid
scaffold (see Figure~\ref{fig:nana}-a).  Of course, a detailed
description of the chemical properties involved is far beyond the
scope of this paper (see also~\cite{braun1998, seeman1999}).

In this work we present \disr{} (Distributed Segment Routing), a distributed
approach focused on addressing the three main challenges that this new fabrication process introduces
in nanoscale Network-on-Chip Design: \emph{(i) limited node complexity}, \emph{(ii) large scale
randomness} and \emph{(iii) high defect rates}.  
The \emph{limited node complexity} aspect is directly
related to the use of complementary DNA tags in order to place circuit
components. Traditional CMOS process introduces complexity 
with larger photolithography mask: more complex (larger) circuits
will only require larger masks. Conversely, self-assembly achieves
complexity by increasing the number of unique DNA tags, since more
different tags means having more control on component placement. Ideally, by specifying a single
and unique tag for each nano device terminal, we could exactly
choose where each component would be placed in the design. But the number of DNA
symbols forming the DNA sequence is limited (sequence of 4 nucleobases
G,A,T,C) and so creating many different tags (of a prefixed length) would
mean make them more similar to each other, increasing the probability
of incorrect/partial matching. To avoid this problem 
the number of unique tags must be limited, thus limiting the
complexity available at each node.
In particular, in this work a budget of about $10,000$ CNFETs per node
has been assumed , as estimated in~\cite{liu_jetcs}.
\emph{Large scale randomness} is the other fundamental condition of
self-assembled technology: DNA tags allow a control placement inside
node grid, but there is no control of where each of these  grids will
be placed on the whole network. As a consequence, other typical
properties of regular networks cannot be guaranteed, e.g. being
connected to a fixed number of neighbors, having a determined
orientation and so on.
Finally, \emph{defect rates}: while they are very hardly tolerated in mask
based top down design, the same nature of a bottom up self-assembly
 process cannot assume such a deterministic device placement
process, thus defect tolerance is something like a design requirement
more than an exception to be avoided.

\begin{figure*}
\centering
\begin{tabular}{cc}
    \includegraphics[width=0.45\textwidth]{pictures/dna2b.eps} &
    \includegraphics[width=0.45\textwidth]{pictures/dna1_complex2.eps} \\
 (a) & (b)
 \end{tabular}
  \caption{(a) Strands of complementary DNA tags self assemble to
  generate larger structures (b) The resulting irregular topology of
  processing(P) and storage(M) elements }
  \label{fig:nana}
\end{figure*}
These aspects of a DNA self-assembly process lead to some important
implications to be addressed when approaching to the design of
a nanoscale Network-on-Chip architecture: computational model should be based on a
distributed network of small processing and storage nodes, randomly
placed and interconnected (see Figure~\ref{fig:nana}-b).
Proof-of-concept of such architectures can be found for
example in~\cite{patwardhan2006_1}, where instruction and data
operands are moved around the network in order to be processed.
In addition, independently of the particular policy chosen to route instruction and
data packets, such networks must be able to guarantee deadlock freedom without any
prior assumption on topology.

In this work, we introduce \disr{}, a Distributed Segment-based
approach to topology discovery and deadlock freedom in large scale DNA
self-assembled on chip networks. Our contribution aims to achieve the
classical properties of a segment-based approach~\cite{mejia_ipdps06}
without requiring any topology graph, external defect map or
centralized algorithm execution.  So the \disr{} approach is not
intended to discover the ``optimal'' segment choice (ideally reachable
with the knowledge of the topology graph) but just to demonstrate a
concrete model that can fit into such complex, irregular and large
sized networks.  It is important to underline that, although we can
assume an architecture similar to the one depicted in
Figure~\ref{fig:nana} as background, the \disr{} topology discovery
approach presented this work is more general and not strictly
dependent on any underlying computational model adopted.

The paper is organized as follows: in Section~\ref{sec:related_works}
we summarize the main approaches for topology agnostic deadlock
freedom. Next, in
Section~\ref{sec:disr_concepts} are described the main elements and
the execution model of the proposed approach. In Section~\ref{sec:simulation},
simulations are carried out with the open-source tool Nanoxim to
demonstrate how \disr{} preserves some of the main segment-based
properties and how compares to the tree-based approaches. Finally a
draft of an architectural implementation is shown in
Section~\ref{sec:implementation} to evaluate the feasibility of the
approach in terms of node complexity.

\section{Background and Contribution}
\label{sec:related_works}
Several works addressed the problem of topology agnostic deadlock
freedom introducing virtual channels~\cite{sancho2002, skeie2002,
skeie2004, koibuchi2003}, but while they show interesting
performances, the low resources available to prevent deadlocks makes them not suitable for
the DNA self-assembled networks we are assuming as
scenario.
Other topology independent algorithms~\cite{schroeder1991,
koibuchi2001, cherkasova1996} do not require any virtual channel but
exploit the concept of ``turn prohibition''.
In particular, authors of~\cite{Patwardhan05evaluatingthe} exploit the creation of a spanning tree of the
topology, placing bidirectional restrictions to avoid that a packet
can traverse a given link in both up and down directions.
The hierarchical nature of this approach can lead to uneven traffic
distribution, with many packets traversing upper links (near to the
root), but this is quite acceptable in classical wide area networks
topologies with a limited number of nodes. Other approaches such as
FX~\cite{sancho2000} mitigate this
issue, but the set of turn restrictions is still prefixed
strictly depending on the particular tree root selected. 
Other solutions try to approach the issue of irregular
topologies by limiting the number~\cite{duato1997, gomez2004, koibuchi2008} or the
location~\cite{zhang2008, flich2008, liu2011} of missing links. This
restriction is clearly unacceptable given the hight-defect rates of
DNA self-assembled networks scenario. For the same reason, we also
avoid considering solutions based on hardware-redundancy to
dynamically recover defects as in~\cite{constantinides2006, kohler2010,  park2006, ebrahimi2013}.

In~\cite{mejia_ipdps06} authors present an approach that solves these
limitations setting turn restrictions locally,
independently from other restrictions. The whole network is
partitioned into segments, and a bidirectional turn
restriction can be freely chosen within a segment in order to guarantee
deadlock freedom while preserving network connectivity. This \emph{locality
independence} property, removing the requirement of choosing a particular tree/root,
would make this approach the best choice for the given scenario;
however, it still requires the knowledge of the
whole network graph in order to find the segments.

The \disr{} approach presented in this work is a first attempt to
achieve the same goals of a segment based turn prohibition without
requiring neither the knowledge of a topology graph nor centralized execution. In particular
\disr{} implements segment discovery and turn prohibition using a
completely different mechanism, based on a distributed exchange of
small setup packets requiring a limited and scalable portion of the available
resource budget. The following main features distinguish \disr{}
from classic approaches to segment partitioning: 
\begin{itemize}
\item No centralized entity is globally responsible and/or aware of
what is going on, i.e. the status of the \disr{} execution is
collectively distributed among the nodes.
\item  No defect map and/or topology graph is used as input,
thus, the topology has to be discovered \emph{while} segments are
created.
\item At the end of the execution, no segment list is created or
stored anywhere: each node is only aware of belonging to a segment,
ignoring the presence of other nodes in the same segment and also the
presence of other segments in the network.
\end{itemize}
